Probing heavy ion collisions using quark and gluon jet substructure

### 2018-03-09 qg_ML.tex
paper version ready for arxiv submission

\section{Quark and Gluon Jet Substructure}
\label{sec:qg}
\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.7\textwidth]{plots/qg_HI}
	   \caption{Illustration of the interplay between jet quenching (vertical) and quark gluon discrimination (horizontal). }
	   \label{fig:quenching_discrimination}
\end{figure}

@@

\begin{figure}[h]
	   \centering
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_jetMass}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_firstRadialMoment}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_halfRadialMoment}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_pTD}
	   \includegraphics[width=0.48\textwidth]{plots/JEWEL_pp_pbpb020_NumberjetConstituents}
	   \caption{Distributions of the jet mass (top left), the girth or the first radial moment (top right), the half radial moment (center left), the $p^{D}_{T}$ (center right) and the pixel multiplicity (bottom) for quark (darker blue) and gluon (lighter orange) jets. The top panels of each individual distribution show the pp ones while the bottom panels correspond to central (0-20\%) PbPb collisions generated by \jw simulations.}
	   \label{fig:jetdistributons_pp_pbpb}
	\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.45\textwidth]{plots/JEWEL_pp_2p76TeV_AverageQuarkJet_Ypt100_jpt50.pdf}
\includegraphics[width=0.45\textwidth]{plots/JEWEL_pp_2p76TeV_AverageGluonJet_Ypt100_jpt50.pdf}
\includegraphics[width=0.45\textwidth]{plots/JEWEL_pbpb020_2p76TeV_AverageQuarkJet_Ypt100_jpt50.pdf}
\includegraphics[width=0.45\textwidth]{plots/JEWEL_pbpb020_2p76TeV_AverageGluonJet_Ypt100_jpt50.pdf}
\caption{Average, pre-processed quark (left) and gluon (right) jet images in the $\eta-\phi$ plane for pp (top) and central PbPb (bottom) collisions generated from \textsc{Jewel}. The color scale represents the average transverse momenta deposited in each pixel in units of GeV/c. }
\label{fig:qgjetimages}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.49\columnwidth]{figures/QCDjetT1_telescoped2.pdf}
\includegraphics[width=.49\columnwidth]{figures/QCDjetT2_telescoped2.pdf}
\includegraphics[width=.49\columnwidth]{figures/QCDjetT3_telescoped2.pdf}
\caption{Illustration of telescoping deconstruction at T1 (top left panel), T2 (top right panel) and T3 (bottom panel) orders for a random pp jet. The crosses are the Winner-Take-All $k_t$ axes. The straight lines are the exclusive subjet boundaries. Particles are sized according to their transverse momenta. See Appendix for more details about the procedures.}
\label{fig:tjetQCD}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Comp_zg_ppcomp_SD_TD2_new.pdf}
	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg_ppcomp_SD_TD2_new.pdf}
	   \caption{Top panels: Distributions of groomed momentum sharing $z_g$ (left) and the T2 order TD subjet momentum fraction $z_{\rm T2}$ (right); Bottom panels: Distributions of groomed jet radius $r_g$ and the T2 order TD angular separation between subjets $\theta_{\rm T2}$ for quark (blue) and gluon (red) jets in pp \jw simulations.}
\label{fig:comp_z_pp_wT2}
\end{figure}

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Comp_z_new.pdf}
	   \caption{Distributions of groomed momentum sharing $z_g$ (left) and TD subjet momentum fraction $z_{\rm T3}$ (right) of quark (blue) and gluon (red) jets in pp (top panels) and central PbPb (bottom panels) collisions in \textsc{Jewel}. Note that $z_{\rm cut}<z_g,z_{\rm T3}<0.5$ and $z_{\rm cut}$ is chosen to be 0.1 in this work. (color online)}
\label{fig:comp_z}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Comp_rg_new.pdf}
	   \caption{Distributions of groomed jet radius $r_g$ (left) and TD angular separation between subjets $\theta_{\rm T3}$ (right) of quark (blue) and gluon (red) jets in pp (top panels) and central PbPb (bottom panels) collisions in \textsc{Jewel}. The quenched $r_{g}$ distribution is significantly modified with enhancement at the edge of jets.}
\label{fig:comp_rg}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Comp_lambda_new.pdf}
	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$ for hard (left) and soft (right panels) subjets with quark (blue dashed) and gluon (red dashed) jets in pp (top) and central PbPb (bottom) collisions in \textsc{Jewel}. }
\label{fig:comp_subjet_m}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Comp_delta_m2.pdf}
	   \caption{Distributions of $\delta m=\sqrt{m_{\rm jet}^2-m_{\rm groomed-jet}^2}$ for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
\label{fig:comp_delta_m2}
\end{figure}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel.pdf}
	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
\label{fig:Lund_full}
\end{figure}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_background.pdf}
	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
\label{fig:Lund_bkg}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{plots/Individual_LundDiagrams_zrel_hardBranch.pdf}
	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
\label{fig:Lund_hard}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_pp_2p76TeV_genLevel_QvsG.pdf}
	\caption{ROC curves of quark v.s. gluon efficiency for particle jets in pp collisions in \textsc{Jewel}. The different colored curves correspond to various classification models studied in this paper, and the quantity in the brackets marked by auc represents the area under the curve. The green solid and dotted lines are for the MLP with physics-motivated variables with and without the pixel multiplicity, respectively. The blue solid line corresponds to the DCNN trained on jet images, while the brown, orange and teal lines are for the TD at T1, T2 and T3 orders, respectively.}
\label{fig:ROC_pp}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_PbPb_2p76TeV_genLevel_QvsG.pdf}
	\caption{ROC curves of quark v.s. gluon efficiency for particle jets in central PbPb (right panel) collisions in \textsc{Jewel}, similar to Figure~\ref{fig:ROC_pp}.}
\label{fig:ROC_pbpb}
\end{figure}

@@

\begin{figure}[h]
	\centering
        \includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_TD_2p76TeV_ppvsPbPb.pdf}
	\caption{ROC curves of quark v.s. gluon efficiency using TD for grid jets in pp (dotted lines) and central PbPb (solid lines) collisions in \textsc{Jewel} with GridSub subtraction. The different colors represent different TD orders (T1: blue, T2: red, T3: green). The corresponding area under the curves are shown in the brackets.}
	\label{fig:ROC_TD_grid}
\end{figure}

\begin{figure}[h]
	\centering
        \includegraphics[width=0.8\textwidth]{plots/JEWELPYTHIA_TD123_2p76TeV_ppvsPbPb_comp_QvsG.pdf}
	\caption{ROC curves of quark v.s. gluon efficiency using TD at the T3 order for jets with pp particle (green dotted), pp grid (blue dotted), PbPb particle (solid green), PbPb grid (solid red) and PbPb grid with GridSub subtraction (solid blue) constituents. The corresponding area under the curves are shown in the brackets.}
	\label{fig:ROC_TD3_grid}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.62\textwidth]{plots/JEWELPYTHIA_2p76TeV_quark_ppvspbpb.pdf}
	   \includegraphics[width=0.62\textwidth]{plots/JEWELPYTHIA_2p76TeV_gluon_ppvspbpb.pdf}
	   \caption{ROC curves of pp-jet v.s. PbPb-jet efficiency for quark (top panel) and gluon (bottom panel) jets with grid discretization. The different colored curves correspond to various classification models studied in this paper.}
\label{fig:ROC_qq_gg}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{figures/topjetT1_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT2_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT3_telescoped2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT1_masses2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT2_masses2.pdf}
\includegraphics[width=.32\columnwidth]{figures/topjetT3_masses2.pdf}
\caption{\label{fig:tjet}(Top row) The TD of a top jet at T1 (left panel), T2 (middle panel) and T3 (right panel) orders. The crosses are the Winner-Take-All $k_t$ axes. The straight lines are the exclusive subjet boundaries. Particles are sized according to their transverse momenta. (Bottom row) The subjet masses at T1 (left panel), T2 (middle panel), and T3 (right panel) orders as the telescoping radius $R_T$ is varied, corresponding to the region of the same color in the top row. The presence of the $W$ in the top jet can be clearly seen at T1 and T2 orders by the plateau of the red line beginning around $R_T=0.3$.}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{figures/QG_invROC.pdf} \includegraphics[width=.32\columnwidth]{figures/W_invROC.pdf} \includegraphics[width=.32\columnwidth]{figures/Top_invROC.pdf}
\caption{\label{ROCs}ROC curves for the DNNs trained on cumulant telescoping deconstruction observables up to T1 through T4 orders and the jet image method using CNNs for quark/gluon discrimination (left panel), boosted $W$ (middle panel) and top (right panel) tagging. The T$N$ performance approximately saturates at T2 (quark/gluon), T1 ($W$ tagging), and T2 (top tagging) orders.}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_top.pdf}
\includegraphics[width=.32\columnwidth]{figures/T1HeavyMass_qcd.pdf}
\includegraphics[width=.32\columnwidth]{figures/T2HeavyMass_qcd.pdf}
\includegraphics[width=.32\columnwidth]{figures/T3HeavyMass_qcd.pdf}
\caption{\label{masses}The heavy subjet mass distributions at T1 (left panels), T2 (middle panels), and T3 (right panels) orders for multiple subjet radii $R_T$ of top jets (top row) and QCD background jets (bottom row). The presence of the $W$ in top jets is evident by the peaks at the $W$ mass at T1 and T2 orders. At T3 order, the heavy subjet mass distributions are QCD-like and are narrower (more quark-like) in top subjets than the wider (more gluon-like) QCD subjets. These features highlight the ability of the telescoping deconstruction to probe subjet substructure.}
\end{figure}

>>>>

\section{Quark and Gluon Jet Substructure}
\label{sec:qg}
\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.7\textwidth]{Fig1}
	   \caption{Illustration of the interplay between jet quenching (vertical) and quark gluon discrimination (horizontal). }
	   \label{fig:quenching_discrimination}
\end{figure}

@@

\begin{figure}[h]
	   \centering
	   \includegraphics[width=0.48\textwidth]{Fig2a}
	   \includegraphics[width=0.48\textwidth]{Fig2b}
	   \includegraphics[width=0.48\textwidth]{Fig2c}
	   \includegraphics[width=0.48\textwidth]{Fig2d}
	   \includegraphics[width=0.48\textwidth]{Fig2e}
	   \caption{Distributions of the jet mass (top left), the girth or the first radial moment (top right), the half radial moment (center left), the $p^{D}_{T}$ (center right) and the pixel multiplicity (bottom) for quark (darker blue) and gluon (lighter orange) jets. The top panels of each individual distribution show the pp ones while the bottom panels correspond to central (0-20\%) PbPb collisions generated by \jw simulations.}
	   \label{fig:jetdistributons_pp_pbpb}
	\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.45\textwidth]{Fig3a}
\includegraphics[width=0.45\textwidth]{Fig3b}
\includegraphics[width=0.45\textwidth]{Fig3c}
\includegraphics[width=0.45\textwidth]{Fig3d}
\caption{Average, pre-processed quark (left) and gluon (right) jet images in the $\eta-\phi$ plane for pp (top) and central PbPb (bottom) collisions generated from \textsc{Jewel}. The color scale represents the average transverse momenta deposited in each pixel in units of GeV/c. }
\label{fig:qgjetimages}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.49\columnwidth]{Fig4a}
\includegraphics[width=.49\columnwidth]{Fig4b}
\includegraphics[width=.49\columnwidth]{Fig4c}
\caption{Illustration of telescoping deconstruction at T1 (top left panel), T2 (top right panel) and T3 (bottom panel) orders for a random pp jet. The crosses are the Winner-Take-All $k_t$ axes. The straight lines are the exclusive subjet boundaries. Particles are sized according to their transverse momenta. See Appendix for more details about the procedures.}
\label{fig:tjetQCD}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{Fig5a}
	   \includegraphics[width=0.9\textwidth]{Fig5b}
	   \caption{Top panels: Distributions of groomed momentum sharing $z_g$ (left) and the T2 order TD subjet momentum fraction $z_{\rm T2}$ (right); Bottom panels: Distributions of groomed jet radius $r_g$ and the T2 order TD angular separation between subjets $\theta_{\rm T2}$ for quark (blue) and gluon (red) jets in pp \jw simulations.}
\label{fig:comp_z_pp_wT2}
\end{figure}

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{Fig6}
	   \caption{Distributions of groomed momentum sharing $z_g$ (left) and TD subjet momentum fraction $z_{\rm T3}$ (right) of quark (blue) and gluon (red) jets in pp (top panels) and central PbPb (bottom panels) collisions in \textsc{Jewel}. Note that $z_{\rm cut}<z_g,z_{\rm T3}<0.5$ and $z_{\rm cut}$ is chosen to be 0.1 in this work. (color online)}
\label{fig:comp_z}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{Fig7}
	   \caption{Distributions of groomed jet radius $r_g$ (left) and TD angular separation between subjets $\theta_{\rm T3}$ (right) of quark (blue) and gluon (red) jets in pp (top panels) and central PbPb (bottom panels) collisions in \textsc{Jewel}. The quenched $r_{g}$ distribution is significantly modified with enhancement at the edge of jets.}
\label{fig:comp_rg}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{Fig8}
	   \caption{Distributions of TD $\lambda$, which is the ratio between subjet mass and subjet $p_T$ for hard (left) and soft (right panels) subjets with quark (blue dashed) and gluon (red dashed) jets in pp (top) and central PbPb (bottom) collisions in \textsc{Jewel}. }
\label{fig:comp_subjet_m}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{Fig9}
	   \caption{Distributions of $\delta m=\sqrt{m_{\rm jet}^2-m_{\rm groomed-jet}^2}$ for quark (blue) and gluon (red) jets in pp (left) and central PbPb (right) collisions in \textsc{Jewel}.}
\label{fig:comp_delta_m2}
\end{figure}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.9\textwidth]{Fig10}
	\caption{Lund diagrams for quark (left) and gluon jets (right) in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
\label{fig:Lund_full}
\end{figure}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.9\textwidth]{Fig11}
	\caption{Lund diagrams for branches removed in the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
\label{fig:Lund_bkg}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.9\textwidth]{Fig12}
	   \caption{Lund diagrams for branches in the hard branch after the soft-drop procedure with $z<z_{\rm cut}=0.1$ for quark (left) and gluon (right) jets in pp (top) and central PbPb collisions in \textsc{Jewel}. See text for the definition of the axes.}
\label{fig:Lund_hard}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{Fig13}
	\caption{ROC curves of quark v.s. gluon efficiency for particle jets in pp collisions in \textsc{Jewel}. The different colored curves correspond to various classification models studied in this paper, and the quantity in the brackets marked by auc represents the area under the curve. The green solid and dotted lines are for the MLP with physics-motivated variables with and without the pixel multiplicity, respectively. The blue solid line corresponds to the DCNN trained on jet images, while the brown, orange and teal lines are for the TD at T1, T2 and T3 orders, respectively.}
\label{fig:ROC_pp}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{Fig14}
	\caption{ROC curves of quark v.s. gluon efficiency for particle jets in central PbPb (right panel) collisions in \textsc{Jewel}, similar to Figure~\ref{fig:ROC_pp}.}
\label{fig:ROC_pbpb}
\end{figure}

@@

\begin{figure}[h]
	\centering
        \includegraphics[width=0.8\textwidth]{Fig15}
	\caption{ROC curves of quark v.s. gluon efficiency using TD for grid jets in pp (dotted lines) and central PbPb (solid lines) collisions in \textsc{Jewel} with GridSub subtraction. The different colors represent different TD orders (T1: blue, T2: red, T3: green). The corresponding area under the curves are shown in the brackets.}
	\label{fig:ROC_TD_grid}
\end{figure}

\begin{figure}[h]
	\centering
        \includegraphics[width=0.8\textwidth]{Fig16}
	\caption{ROC curves of quark v.s. gluon efficiency using TD at the T3 order for jets with pp particle (green dotted), pp grid (blue dotted), PbPb particle (solid green), PbPb grid (solid red) and PbPb grid with GridSub subtraction (solid blue) constituents. The corresponding area under the curves are shown in the brackets.}
	\label{fig:ROC_TD3_grid}
\end{figure}

@@

\begin{figure}[t]
	   \centering
	   \includegraphics[width=0.62\textwidth]{Fig17a}
	   \includegraphics[width=0.62\textwidth]{Fig17b}
	   \caption{ROC curves of pp-jet v.s. PbPb-jet efficiency for quark (top panel) and gluon (bottom panel) jets with grid discretization. The different colored curves correspond to various classification models studied in this paper.}
\label{fig:ROC_qq_gg}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{Fig18a}
\includegraphics[width=.32\columnwidth]{Fig18b}
\includegraphics[width=.32\columnwidth]{Fig18c}
\includegraphics[width=.32\columnwidth]{Fig18d}
\includegraphics[width=.32\columnwidth]{Fig18e}
\includegraphics[width=.32\columnwidth]{Fig18f}
\caption{\label{fig:tjet}(Top row) The TD of a top jet at T1 (left panel), T2 (middle panel) and T3 (right panel) orders. The crosses are the Winner-Take-All $k_t$ axes. The straight lines are the exclusive subjet boundaries. Particles are sized according to their transverse momenta. (Bottom row) The subjet masses at T1 (left panel), T2 (middle panel), and T3 (right panel) orders as the telescoping radius $R_T$ is varied, corresponding to the region of the same color in the top row. The presence of the $W$ in the top jet can be clearly seen at T1 and T2 orders by the plateau of the red line beginning around $R_T=0.3$.}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{Fig19a}
\includegraphics[width=.32\columnwidth]{Fig19b}
\includegraphics[width=.32\columnwidth]{Fig19c}
\caption{\label{ROCs}ROC curves for the DNNs trained on cumulant telescoping deconstruction observables up to T1 through T4 orders and the jet image method using CNNs for quark/gluon discrimination (left panel), boosted $W$ (middle panel) and top (right panel) tagging. The T$N$ performance approximately saturates at T2 (quark/gluon), T1 ($W$ tagging), and T2 (top tagging) orders.}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=.32\columnwidth]{Fig20a}
\includegraphics[width=.32\columnwidth]{Fig20b}
\includegraphics[width=.32\columnwidth]{Fig20c}
\includegraphics[width=.32\columnwidth]{Fig20d}
\includegraphics[width=.32\columnwidth]{Fig20e}
\includegraphics[width=.32\columnwidth]{Fig20f}
\caption{\label{masses}The heavy subjet mass distributions at T1 (left panels), T2 (middle panels), and T3 (right panels) orders for multiple subjet radii $R_T$ of top jets (top row) and QCD background jets (bottom row). The presence of the $W$ in top jets is evident by the peaks at the $W$ mass at T1 and T2 orders. At T3 order, the heavy subjet mass distributions are QCD-like and are narrower (more quark-like) in top subjets than the wider (more gluon-like) QCD subjets. These features highlight the ability of the telescoping deconstruction to probe subjet substructure.}
\end{figure}
